<h2> <a name="1_Introduction" style="pointer-events: none; cursor: default;" ><font color="black"> 1. Introduction </font></a> </h2>



This report presents a statistical analysis of the wine quality dataset released by the Portuguese firm Vinho Verde, aiming at classifying sample qualities based on their physical-chemical properties. To this end, various models and different approaches are compared so to individuate the best in terms of out-of-sample predictions.

The original data are split according to the wine typology, with  4898 white wine samples and 1599 are reds. However, the two sets are considered altogether in this report, and a variable *type* is added with value 1 for white wines and 0 otherwise.  All in all, the final dataset is then composed of a total of $n=6497$ statistical units, with 11 variables expressing physical-chemical qualities and the additional feature *type*. 
The target variable is *quality* and measures the sample quality as an ordinal categorical variable in the range from 1 to 10. Each of its values is obtained as the median of at least 3 votes expressed by expert sommeliers.

```{r, fig.width=9.5, fig.height=5, echo=FALSE}
file = "../results/EXPLORATORY_ANALYSES/quality_distr.Rdata"
load( file )
quality_distr
```

<br></br>
<br></br>

As we can see, the actual range of the target is the inteval $[3,9] \in \mathbb{N}$. 
Also, the categories 5 and 6 are the most numerous and they alone account for the 76.6% of the whole sample.  This may be a problem when training since the model could learn to predict just the most represented classes.
For this reason, we try to balance the target variable by considering a binary version obtained as follows:  

$$ \begin{equation}
  \textit{binary_quality} = 
  \begin{cases}
   1    &\textit{se quality} > 5  \\ 0       &\textit{altrimenti}
   \end{cases}
\end{equation}$$


The physical-chemical information available is contained in the following variables:

* Fixed Acidity  (*fixed.acidity*)
* Volatile Acidity (*volatile.acidity*)
* Citric Acid (*citric.acid*)
* Residual Sugar (*residual.sugar*)
* Chlorides (*chlorides*)
* Free Sulphur Dioxide (*free.sulfur.dioxide*)
* Total Sulphur Dioxide (*total.sulfur.dioxide*)
* Density (*density*)
* pH (*pH*)
* Sulphates (*sulphates*)
* Alcohol (*alcohol*)

As first insight, the following plots illustrate the distributions of each variable in both subsamples (red and yellow curves) and in the whole dataset (blue curve):

<br></br>
```{r, fig.width=9.5, fig.height=8, echo=FALSE}
file = "../results/EXPLORATORY_ANALYSES/variables_density.Rdata"
load( file )


p1 = plots[[1]]
p2 = plots[[2]]
p3 = plots[[3]]
p4 = plots[[4]]
p5 = plots[[5]]
p6 = plots[[6]]
p7 = plots[[7]]
p8 = plots[[8]]
p9 = plots[[9]]
p10 = plots[[10]]
p11 = plots[[11]]
p12 = plots[[12]]

multi_p = multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols = 4 )



```

<br></br>
<br></br>


The features are mostly *bell-shaped*, except for *total.sulfur.dioxide*, *citric.acid* and *chlorides* that show some symptoms of being multimodal. Moreover, all the variables have a quite similar range of variation, thus suggesting some pre-processing (e.g. standardisation). For this reason, we performed the analysis both with and without preliminary transformations of the original data, and we found non-significant differences between the two approaches. Hence, only results obtained without pre-processing are shown in the following for convenience.


Although correlation does not mean causation, it makes sense to explore the dependency structure between the target variable and each of the predictors.

For this purpose, the correlation matrix computed on the whole dataset is reported below:

```{r, fig.width=9.5, fig.height=4, echo=FALSE}

file = "../results/EXPLORATORY_ANALYSES/corrplot.Rdata"
load( file )
corrplot

```

<br></br>
<br></br>

The target quality (in its original format) is very poorly correlated with the predictors, and just in two cases we observe a linear correlation coefficient above 0.3 (*alcohol* and *density*).
On the other hand,  predictors appear highly linked to one another. In particular, the strongest relationships we can observe are:

- *free.sulfur.dioxide* - *total.sulfur.dioxide* (0.72)
- *density* - *alcohol* (-0.687)
- *density* - *residual.sugar* (0.553)

However, given the categorical nature of the variable Quality, it would be more appropriate to study its correlation structure based on the Spearman coefficient rather than the Pearson one:

```{r, fig.width=9.5, fig.height=4, echo=FALSE}
file = "../results/EXPLORATORY_ANALYSES/corr_Y.Rdata"
load( file )
corr_Y
```

<br></br>
<br></br>


The Spearman coefficient, $\rho$, measures the degree of relation between two ordinal categorical variables.
Unlike the Pearson coefficient, Spearman $\rho$ assumes just a monotonic relationship between the variables (not necessarily a linear one).

For more details on the dataset check:

https://archive.ics.uci.edu/ml/datasets/wine+quality


The analyses presented in the following are conducted considering about 60% of the data as training set and the remaining 40% for testing.


<br></br>
<br></br>
